---
title: "Additional carat features"
author: "Tejus"
date: "31/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2); library(caret); library(gridExtra);
library(ISLR)
library(reshape2)
library(Hmisc)
```

## Plotting predictors

* `iris` data set gives the measurements in centimeters of the variables sepal length and width and petal length and width, respectively, for 50 flowers from each of 3 species of iris. The species are Iris setosa, versicolor, and virginica.

```{r load_wage}
data("iris")
summary(iris)
```

* Build a model 

```{r training, echo=FALSE}

Train_Index <- createDataPartition(y = iris$Species, p =0.7, list = FALSE)
train <- iris[Train_Index,]
test <- iris[-Train_Index,]
      
```

### Scatterplot Matrix with *featurePlot* (*caret* package)

#### Scatterplot matrix and factor variables

* Scatterplot with pairs of data

```{r featurePlot, out.width= "120%", out.height="90%"}

featurePlot(x=train[,c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")],
            y = train$Species,
            plot="pairs", 
            ## Add a key at the top 
            auto.key = list(columns = 3))

```


* Scatterplot with pairs of data and ellipses


```{r featurePlot_ellipse, out.width= "120%", out.height="90%"}

featurePlot(x=train[,c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")],
            y = train$Species,
            plot="ellipse", 
            ## Add a key at the top 
            auto.key = list(columns = 3))

```


* Overlayed Density Plots

```{r featurePlot_density, out.width= "100%"}

featurePlot(x=train[,c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")],
            y = train$Species,
            plot="density", pch = "|", 
            layout = c(4, 1), adjust = 1.5, 
            scales = list(y = list(relation="free"),
                          x = list(relation="free")),
            ## Add a key at the top 
            auto.key = list(columns = 3))

```

* Box Plots

```{r featurePlot_box, out.width= "100%", out.height="40%"}

featurePlot(x=train[,c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")],
            y = train$Species,
            plot="box", 
            layout = c(4, 1), 
            scales = list(y = list(relation="free"),
                          x = list(rot=90)),
            ## Add a key at the top 
            auto.key = list(columns = 3))

```


#### Scatterplot matrix and continous variables

* For regression, the Boston Housing data is used :

* Housing data for 506 census tracts of Boston from the 1970 census. The dataframe `BostonHousing2` the corrected version with additional spatial information.

* The original data are 506 observations on 14 variables, medv (median value of owner-occupied homes in USD 1000's) being the target variable :

```{r BostonHousing, cache=TRUE}

library(mlbench)
data(BostonHousing2)
regVar <- c("age", "lstat", "tax")
str(BostonHousing2[, regVar])

```

* `featurePlot` can be used to create scatter plots of each of the predictors with the outcome.

```{r scatter, out.width="60%"}

theme1 <- trellis.par.get()
theme1$plot.symbol$col = rgb(.2, .2, .2, .4)
theme1$plot.symbol$pch = 16
theme1$plot.line$col = rgb(1, 0, 0, .7)
theme1$plot.line$lwd <- 2
trellis.par.set(theme1)

# actual plot
featurePlot(x = BostonHousing2[, regVar], 
            y = BostonHousing2$medv, 
            plot = "scatter",
            type = c("p", "smooth"),
            span = .5,
            layout = c(3, 1))

```


### Ggplot2 is always handy with smoothing

```{r }

Boston_Melted <- melt(BostonHousing2[,c(regVar, "medv")], measure.vars = regVar )

g <- ggplot(Boston_Melted, aes(x = value, y = medv)) + geom_point(aes(color = variable)) + facet_grid(.~variable, scales = "free_x") + geom_smooth(method='lm')
g
```



### cut2, making factors (*Hmisc* package)

* Adding factors to variables

```{r cut2,fig.height=4,fig.width=6,cache=TRUE, warning=FALSE}

cutAge <- cut2(BostonHousing2$age,g=3)
cutLstat <- cut2(BostonHousing2$lstat,g=3)
cutTax <- cut2(BostonHousing2$tax,g=3)

ExtraFactors <- c(cutAge,  cutLstat, cutTax)

table(cutAge)

```



### Ggplot for boxplot 

* Boxplot with additional factors in our Regression variables

```{r cutPlots, out.width = "120%" }
g1 <- qplot(cutAge, medv, data = BostonHousing2, fill = cutAge, geom = c("boxplot"))
g2 <- qplot(cutTax, medv, data = BostonHousing2, fill = cutTax, geom = c("boxplot"))
g3 <- qplot(cutLstat, medv, data = BostonHousing2, fill = cutLstat, geom = c("boxplot"))

grid.arrange(g1, g2, g3, ncol = 2)
```






