---
title: "Forecasting"
author: "Tejus"
date: "15/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(quantmod); library(forecast)
```

### What is different?

* Data are dependent over time
* Specific pattern types
  * Trends - long term increase or decrease
  * Seasonal patterns - patterns related to time of week, month, year, etc.
  * Cycles - patterns that rise and fall periodically
* Subsampling into training/test is more complicated
* Similar issues arise in spatial data 
  * Dependency between nearby observations
  * Location specific effects
* Typically goal is to predict one or more observations into the future. 
* All standard predictions can be used (with caution!)

#### Things to remember

* Beware spurious correlations!

* Beware extrapolation!


### Forecasting with Yahoo data

```{r loadGOOG, cache=TRUE}

from.dat <- as.Date("01/01/08", format="%m/%d/%y")
to.dat <- as.Date("12/31/13", format="%m/%d/%y")
getSymbols("GOOG", src="yahoo", from = from.dat, to = to.dat)
head(GOOG)
```

* Summarise 

```{r TimeSeries, cache=TRUE}
mGoog <- to.monthly(GOOG)
googOpen <- Op(mGoog)
ts1 <- ts(googOpen,frequency=12)
plot(ts1,xlab="Years+1", ylab="GOOG")

```

#### Examples of time series decomposition

* __Trend__  - Consistently increasing pattern over time 
* __Seasonal__ -  When there is a pattern over a fixed period of time that recurs.
* __Cyclic__ -  When data rises and falls over non fixed periods

[Otexts](https://www.otexts.org/fpp/6/1)


```{r Decompose, cache=TRUE}
plot(decompose(ts1),xlab="Years+1")
```

#### Training and test sets

* Have to build sets with consecutive time points

```{r trainingTest,dependson="tseries",fig.height=4.5,fig.width=4.5}
ts1Train <- window(ts1,start=1,end=5)
ts1Test <- window(ts1,start=5,end=(7-0.01))
ts1Train
```

### Forecasting with simple moving average

$$ Y_{t}=\frac{1}{2*k+1}\sum_{j=-k}^k {y_{t+j}}$$

```{r Algs,fig.height=4.5,fig.width=4.5}
plot(ts1Train)
lines(ma(ts1Train,order=3),col="red")
```

* Prediction will be an average of the previous time points out to a particular time.

### Exponential smoothing

* Weight nearby time points with higher values than those that are far.

__Example - simple exponential smoothing__
$$\hat{y}_{t+1} = \alpha y_t + (1-\alpha)\hat{y}_{t-1}$$

* Different classes to choose from

<img class=center src=./Assets/expsmooth.png height=300>

```{r SES_Eg}
ets1 <- ets(ts1Train,model="MMM")
fcast <- forecast(ets1)
# You get prediction bounds also
plot(fcast); lines(ts1Test,col="red")

## For Accuracy 
accuracy(fcast,ts1Test)

```


### Notes and further resources

* [Forecasting and timeseries prediction](http://en.wikipedia.org/wiki/Forecasting) is an entire field
* Rob Hyndman's [Forecasting: principles and practice](https://www.otexts.org/fpp/) is a good place to start
* Cautions
  * Be wary of spurious correlations
  * Be careful how far you predict (extrapolation)
  * Be wary of dependencies over time
* See [quantmod](http://cran.r-project.org/web/packages/quantmod/quantmod.pdf) or [quandl](http://www.quandl.com/help/packages/r) packages for finance-related problems.


