---
title: "Prediction with Regression"
author: "Tejus"
date: "13/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret);data(faithful); set.seed(333)
library(ISLR); library(ggplot2);data(Wage); 

```

## Example: Old faithful eruptions

```{r faithful}
summary(faithful)

# fit a model
inTrain <- createDataPartition(y = faithful$waiting, p = 0.5, list = FALSE)
training <- faithful[inTrain,]; testing <- faithful[-inTrain,]

```

## Fit a linear model 

$$ ED_i = b_0 + b_1 WT_i + e_i $$

```{r faithlm,dependson="faith",fig.height=4,fig.width=4}
lm1 <- lm(eruptions ~ waiting,data=training)
summary(lm1)
```

* Prediction

```{r lm_predict}
newdata <- data.frame(waiting=80)
predict(lm1,newdata)
```

## Plot predictions - training and test

```{r ,dependson="faithlm",fig.height=4,fig.width=8}
par(mfrow=c(1,2))
plot(training$waiting,training$eruptions,pch=19,col="blue",xlab="Waiting",ylab="Duration")
lines(training$waiting,predict(lm1),lwd=3)
plot(testing$waiting,testing$eruptions,pch=19,col="blue",xlab="Waiting",ylab="Duration")
lines(testing$waiting,predict(lm1,newdata=testing),lwd=3)
```

* For test the fit is shifted near `0` little bit, but that's expected.

## Prediction intervals

* Prediction intervals

```{r fig.height=4,fig.width=4}
pred1 <- predict(lm1,newdata=testing,interval="prediction")
ord <- order(testing$waiting)
plot(testing$waiting,testing$eruptions,pch=19,col="blue")
matlines(testing$waiting[ord],pred1[ord,],type="l",col=c(1,2,2),lty = c(1,1,1), lwd=3)
```


* With `caret`

```{r caretfaith,dependson="faith",fig.height=4,fig.width=4}
modFit <- train(eruptions ~ waiting,data=training,method="lm")
summary(modFit$finalModel)
```


## Predicting with Regression Multiple Covariates

* Wage data

```{r loadData,cache=TRUE}
Wage <- subset(Wage,select=-c(logwage))
summary(Wage)

inTrain <- createDataPartition(y=Wage$wage, p=0.7, list=FALSE)
training <- Wage[inTrain,]; testing <- Wage[-inTrain,]
dim(training); dim(testing)
```


* In the plot, age versus wage : Some outlir trend is seen at top

```{r fig.height=4,fig.width=6, cache=TRUE}
qplot(age,wage,data=training)
```


* When the same plot, age versus wage is colored by jobclass there is some explanation seen


```{r fig.height=4,fig.width=6, cache=TRUE}
qplot(age,wage,colour=jobclass,data=training)
```


* Same with when colored by education. 


```{r fig.height=4,fig.width=6, cache=TRUE}
qplot(age,wage,colour=education,data=training)
```

* Seems like the variables `jobClass & Education` are correlated.


## Fit a linear model 

$$ ED_i = b_0 + b_1 age + b_2 I(Jobclass_i="Information") + \sum_{k=1}^4 \gamma_k I(education_i= level k) $$

```{r modelFit, cache=TRUE,fig.height=4,fig.width=4}
modFit<- train(wage ~ age + jobclass + education,
               method = "lm",data=training)
finMod <- modFit$finalModel
print(modFit)
```

* Color by variables not used in the model 

```{r,dependson="modelFit",fig.height=4,fig.width=6}
qplot(finMod$fitted,finMod$residuals,colour=race,data=training)
```




